Carcinoma

### 2022-05-16 inst/carcinoma.Rmd
minor rewriting

@@

\begin{equation}\tag{1}
i^+,j^-\frac{p_i}{p_i+p_j},\qquad i^-,j^+\frac{p_j}{p_i+p_j}
\end{equation}

Operationally the way we implement this in `hyper2` idiom is to posit
an (unobservable) ranking on the clinicians: we place the clinicians
in order from most convinced of the presence of carcinoma to the least
convinced.  Thus a ranking of `DACBEGF` means that clinician `D` was
most convinced of the presense of carcinoma and `F` least convinced.
Given this, and given a single slide, our actual observation might be
that `D` and `A` diagnosed presence, and the others diagosed absence,
of carcinoma.  The corresponding entry would be

@@

for(i in seq_len(nrow(carcinoma_table))){
      jj <- unlist(carcinoma_table[i,1:7])
      negative <- names(jj)[jj==0]
      positive <- names(jj)[jj==1]
      print(negative)
      print(positive)
      Wi[[i]] <- ggrl(W,positive,negative)
}
carcinoma <- lsl(Wi,powers= carcinoma_table$n)
```

@@

```{r,label=countcarcinoma,cache=TRUE}
carcinoma_count <- colSums(sweep(carcinoma_table[,1:7],1,carcinoma_table$n,"*"))
names(carcinoma_count) <- colnames(carcinoma_table)[1:7]
carcinoma_count <- carcinoma_count/sum(carcinoma_count)
carcinoma_count
system.time(lc <- loglik_lsl(carcinoma_count,carcinoma))
```

@@

```{r,label=calccloglikzipf,cache=TRUE}
z <- zipf(7)
names(z) <- LETTERS[1:7]
system.time(lz <- loglik_lsl(z,carcinoma))
```

@@

```{r,label=calcloglikequalp,cache=TRUE}
e <- equalp(W)
system.time(le <- loglik_lsl(e,carcinoma))
```

@@

```{r,label=calculateloglikformaxp, cache=TRUE}
carcinoma_maxp <- 
c(A = 0.118778759012282, B = 0.542873969223425, C = 0.00654044305270073, 
D = 0.00242693714895262, E = 0.202821434317779, F = 0.00130596252816986, 
G = 0.125252494716691)
system.time(lmax <- loglik_lsl(carcinoma_maxp,carcinoma))
```

We are now in a position to compare  these three points:

```{r,label=showthreemethods}
results <- c(count=lc, zipf=lz, equal=le,maxp=lmax)
results <- results - max(results)
results
```

>>>>

(takes maybe twenty minutes to run without cache.  The value of
`carcinoma_maxp` is calculated separately [discussed below], and that
takes a few days to run).

@@

\begin{equation}\tag{1}
\operatorname{Prob}(i^+,j^-) = \frac{p_i}{p_i+p_j},\qquad
\operatorname{Prob}(i^-,j^+) = \frac{p_j}{p_i+p_j}
\end{equation}

(here superscripts $+$ and $-$ refer to a positive and negative
diagnosis respectively).  Operationally the way we implement this in
`hyper2` idiom is to posit an (unobservable) ranking on the
clinicians: we place the clinicians in order from most convinced of
the presence of carcinoma to the least convinced.  Thus a ranking of
`DACBEGF` means that clinician `D` was most convinced of the presense
of carcinoma and `F` least convinced.  Given this, and given a single
slide, our actual observation might be that `D` and `A` diagnosed
presence, and the others diagosed absence, of carcinoma.  The
corresponding entry would be

@@

for(i in seq_len(nrow(carcinoma_table))){
      jj <- unlist(carcinoma_table[i,1:7])
      negative <- names(jj)[jj==0]
      positive <- names(jj)[jj==1]
      Wi[[i]] <- ggrl(W,positive,negative)
}
carcinoma <- lsl(Wi,powers= carcinoma_table$n)
```

@@

```{r,label=countcarcinoma,cache=TRUE}
carcinoma_count <- colSums(sweep(carcinoma_table[,1:7],1,carcinoma_table$n,"*"))
names(carcinoma_count) <- colnames(carcinoma_table)[1:7]
carcinoma_count <- carcinoma_count/sum(carcinoma_count)
carcinoma_count
lc <- loglik_lsl(carcinoma_count,carcinoma)
```

@@

```{r,label=calccloglikzipf,cache=TRUE}
z <- zipf(7)
names(z) <- LETTERS[1:7]
z
lz <- loglik_lsl(z,carcinoma)  # takes about 30 seconds to run
```

@@

```{r,label=calcloglikequalp,cache=TRUE}
e <- equalp(W)
e
le <- loglik_lsl(e,carcinoma) # takes about 30 seconds to run
```

@@

```{r,label=calculateloglikformaxp, cache=TRUE}
carcinoma_maxp <- 
c(A = 0.118778759012282, B = 0.542873969223425, C = 0.00654044305270073, 
D = 0.00242693714895262, E = 0.202821434317779, F = 0.00130596252816986, 
G = 0.125252494716691)
carcinoma_maxp
lmax <- loglik_lsl(carcinoma_maxp,carcinoma)
```

[of course, if `carcinoma_maxp` is available, we would use this as a
starting point, viz
`maxp_lsl(carcinoma,startp=indep(carcinoma_maxp))`].  With this, we
are now in a position to compare these four points:

```{r,label=showthreemethods}
(results <- c(count=lc, zipf=lz, equal=le,maxp=lmax))
results <- results - max(results)
results
```

We can use Wilks here to assess the null of `carcinoma_count`.  We
have $-2\log(\Lambda/\Lambda_0)\sim\chi^2_6$, giving us a $p$-value of

```{r,label=calculatechisquaredpvalueforcount}
pchisq(-2*results[1],df=6,lower.tail=FALSE)
```

What is the difference between `carcinoma_count` and `carcinoma_maxp`?

```{r}
carcinoma_count
carcinoma_maxp
ordertransplot(carcinoma_count,carcinoma_maxp)
ordertransplot(log(carcinoma_count),log(carcinoma_maxp),plotlims=c(-8,0))
```
